Multi-resolution Neural Field Estimation

### 2011-03-08 Multiresolution_framework/ltx/manuscript.tex
edit to result section

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right). 
	\, dr'dr}
\end{align}
Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multipling by $\boldsymbol\mu\left(\mathbf r\right)$, and integrating over the space gives
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I,
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed, white noise process with covariance \parham{(see \cite{Scerri2009} for proof)}
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

@@

\section{Results}
To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using \eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation}, enabling a comparison between true and estimated parameters. 200 realizations of 1~s of data were generated, where the estimation was applied to the final 900~ms, allowing the model's initial transients to die out. The observation noise was set to $\boldsymbol\Sigma_{\epsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance was set to $\eta(r-r') = \phi_{3,-2}(r-r')$ (B-spline, $j=3$). The sampling period and the synaptic time constant were set to $T_s = 1$~ms and $\tau = 10$~ms, respectively. The firing rate slop, $\varsigma$ was set to $0.56~mV^{-1}$. The distance between adjacent sensors was $0.5$~mm, resulting into $n_y = 161$ observations. The observation kernel was modeled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude. The spacing and bandwidth of the sensors allowed for the full spatial bandwidth of the field to be observed.  %\ken{if we need to gain space I would remove the rest of this paragraph and put all parameters in a table} 
The spatial cutoff frequency of the observed field, $y_t(r_n)$, was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of 5, the cutoff spatial frequency was $\nu_{cy} \approx 6.9 $ cycles/mm, \parham{This was obtained by averaging the power in the spatial frequency (over time) of the spatial frequency from the observations} \cite{Freestone2011}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = 131$ states.

>>>>

To derive the state-space model, we first define the matrices 
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right). 
	\, dr'dr}
\end{align}
Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multipling by $\boldsymbol\mu\left(\mathbf r\right)$, and integrating over the space gives
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I,
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

@@

\section{Results}
To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using \eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation}, enabling a comparison between true and estimated parameters. 200 realizations of 1~s of data were generated, where the estimation was applied to the final 900~ms, allowing the model's initial transients to die out. The observation noise was set to $\boldsymbol\Sigma_{\epsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance was set to $\eta(r-r') = \phi_{3,-2}(r-r')$ (B-spline, $j=3$). The sampling period and the synaptic time constant were set to $T_s = 1$~ms and $\tau = 10$~ms, respectively. The firing rate slop, $\varsigma$ was set to $0.56~mV^{-1}$. The distance between adjacent sensors was $0.5$~mm, resulting into $n_y = 161$ observations. The observation kernel was modeled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude. The spacing and bandwidth of the sensors allowed for the full spatial bandwidth of the field to be observed.  %\ken{if we need to gain space I would remove the rest of this paragraph and put all parameters in a table} 
The spatial cutoff frequency of the observed field, $y_t(r_n)$, was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of 5, the cutoff spatial frequency was $\nu_{cy} \approx 6.9 $ cycles/mm, \parham{This was obtained by averaging the spectral power of the spatial frequency (over time) from the observations} \cite{Scerri2009}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = 131$ states.
